\LARGE \bf How Children Perceive and Interact with a Robot that Behaves Unexpectedly - The Domino Experiment

### 2014-09-04 paper.tex
Created a new, 'clean' paper. Started to add relevant parts

\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{I.2.9}{Artificial Intelligence}{Robotics}[commercial robots and applications]

%\terms{Theory}

\keywords{anthropomorphism; children; engagement; human-robot interaction; mental models of robot behavior; wizard-of-oz}

\section{Introduction}

In the previous study, we investigated how children interact and relate to
Ranger. We found that manipulating the robot's behavior (\textit{proactive vs.
reactive}) can result in a different interaction style with the robot
(\textit{playful vs. task-oriented}). Generally, the robot was able to motivate
children to tidy up by putting their toys into the box.

However, the observed engagement and motivation may be due to an initial
fascination for the robot which is likely to fade out over time. Sustaining
interaction beyond novelty in HRI is challenging, and in child-robot
interaction, in particular. Also in the Roomba Study, we saw that initially,
children were fascinated by Roomba and played with it (even though this robot is
not made to encourage children to interact with it) but most of the children
lost their fascination soon and engaged less in using or playing with the robot.
Previous work comes to a similar conclusion. For instance, a study with the Pleo
robot in family homes showed that children interacted less with the robot after
some time \cite{fernaeus_how_2010}. We can expect a similar trend with Ranger.
This makes it important to explore possibilities of how to sustain children's
interaction with Ranger. The question is, how we can enhance children's interest
in Ranger so that they still engage with it after some time, and still feel
motivated by the robot to tidy up. Can we achieve this without increasing the
complexity and cost of the robot too much, keeping in mind the affordability of
the robot and its prospective deployment in a domestic environment?

In this chapter, we present a study that explores possibilities of sustaining
children's engagement with Ranger by manipulating the robot's behavior in such
way that it appears \textit{unexpected} to the children. We examine how
different variations of robot behavior impact children's interaction with Ranger
and their perception of it (\eg in terms of attributing intention and cognitive
abilities to the robot).

\section{Scope and Research Goals}

@@

\section{Related Work}

\subsection{Motivation and Contributions}

The overall motivation of this study is to explore ways of fostering engagement and sustaining children's interest in the robot. Ultimately, the robotic box Ranger aims to motivate children to tidy up, and engagement with the robot is a prerequisite for this: \textit{``Engagement is crucial, because it is typically a prerequisite for other system objectives: If a user stops interaction with a system, then it cannot have any further impact''} \cite[p.~648]{bickmore_maintaining_2010}.

The question is how can we extend children's interest in the robot and sustain their engagement? This question is related to the \textit{novelty effect} which is commonly observed in human-robot interaction. Engagement beyond novelty is challenging to achieve, and optimally a long-term study should be used to investigate the interaction over time. With the current Ranger prototype which is operated by a human Wizard, a longitudinal study in a natural setting was not feasible at this point of time.\footnote{While we were planning this study, a second version of the Ranger prototype was under development. However, we wanted to take any opportunity to study children's interaction with the Ranger in more detail, and therefore set up another short-term study that would focus on a specific aspect of the interaction.} Although we see it as a limitation, most of the HRI studies that investigated engagement and attributions of intentionality to a robot were short-term studies but nevertheless produced relevant results that can encourage future research. Also the Domino Study is a short-term investigation and does not directly look at long-term usage but rather addresses the question of engagement with the robot, which in turn is a pre-requisite for long-term engagement.\\

We studied the effect of three different types of unexpected robot behavior (sometimes also labeled as \textit{misbehavior}) by means of a short-term interaction study in a controlled lab-setting.

If we are able to show that the robot's variation of behavior can improve short-term engagement, it is likely that this also leads to improvement in long-term engagement \cite{bickmore_maintaining_2010}.

\begin{figure}[!h] 
    \centering 
    \includegraphics[width=0.5\columnwidth]{domino_header.png} 
    \caption[How do Children Interact with a Robot that Behaves
    Unexpectedly?]{\small \textbf{How do children perceive and interact with a
    robot that behaves unexpectedly?} In an interaction study in the lab, 26
    children aged 4-5 years played with a robot that behaved unexpectedly from time
    to time. We studied children's engagement with the robot and their attributions
    of intentionality to it.}

    \label{fig:domino-header} 
\end{figure}

@@

In the Domino Study, we analyze child-robot interaction with a robot that shows
unexpected behavior. In a playful scenario which was set up in a laboratory
environment (Figure~\ref{fig:domino-setup}), 26 children aged 4-5 years were
assembling a domino game together. Each group consisted of two children and the
Ranger robot, which was used to transport domino tiles between the two children.

Ranger usually behaved correctly (expected behavior), coming over to a child
after being called and delivering the domino tile to the other child. However,
in pre-defined rounds Ranger showed unexpected behavior, when a child called the
robot. We defined three different types of \textit{misbehavior} that were tested
in a between-subjects study design:

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
        over, the robot goes wrong but recognizes its mistake and repairs. We
        expect this to be perceived as \textit{``to err is human''}, and assume
        increased attributions of human-likeness to the robot.\footnote{The
        attribution of human-likeness to a robot is labeled as
        \textit{anthropomorphism}, and we will hereafter sometimes use ``attribute
        intention'' and ``anthropomorphize'' synonymously.}

    \item The robot gets \textbf{lost}: When called by the child to come over,
        the robot goes wrong, without any observable reason, and remains at the
        wrong location. We expect this to be perceived as a bug or system error
        which causes the robot to not work correctly, and assume few
        attributions of human-likeness to the robot.

    \item The robot \textbf{disobey}: When called by the child to come over, the
        robot shows the intention to not obey by literally ``shaking its head'',
        which should reflect the negative reply to the child's request. The
        robot then goes to a wrong location and remains there while it continues
        to shake its head (this cue is not used in the lost behavior). We expect
        the disobey behavior to be perceived as the robot having an
        \textit{``own will''}, and we assume this leads to increased
        attributions of human-likeness (intention) to the robot.

@@

We analyzed children's reaction focusing on two main aspects. On one hand,
children's \textbf{behavior} (their reactions) toward the unexpected robot
behavior was studied in terms of \textbf{active engagement} with the robot. On
the other hand, we analyzed children's \textbf{perception} of the robot in term
of \textbf{anthropomorphism} -- the attribution of human-like characteristics,
such as cognitive abilities and the ability to show intentions. We assumed that
in general a robot that behaves unexpectedly from time to time can promote
engagement and lead children to attribute intention to it. \cite{short_no_2010}
found that participants anthropomorphize a cheating robot more than a robot that
always behaves fairly, and also evaluate the interaction as more engaging. Based
on the related work we formulate the following two hypotheses:

@@

Our research questions deal with both children's observable behavior and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} cannot only be measured as a specific type of
perception but also as an observable behavior in the interaction itself. Thus,
we would like to take a first step in bringing things together and develop an
\textbf{index of anthropomorphism} that considers both children's perception and
interaction aspects (qualitatively). Previous work suggests that a social
relation to a robot (we view anthropomorphism as a specific type of social
relation) reflects an increased engagement and can be effective in sustaining
interaction. Consequently, we hypothesize a positive correlation between the
anthropomorphic perception of the robot and the amount of engagement in the
interaction.

%\subsection{Engagement}

%\subsection{Unexpected Robot Behavior}

\section{Method and Procedure}

\subsection{Description of the Scenario}

@@

We used a self-made domino consisting of 10 wooden tiles (10~x~20~x~1.5~cm) with
pictures of comic farm animals: cow, sheep, hen, donkey, duck, pig, rabbit. The
pictures were taken from a commercially available domino game adapted to the age
of the children (produced by Djeco, advised age 3+).

@@

\subsection{Manipulation of Robot Behavior}
\label{sec:domino-manipulation}

We used the same prototype as in the previous study, and Ranger was again
controlled by a human Wizard, who was in the same room (see
Figure~\ref{fig:domino-setup}). This was to ensure that the Wizard could see
children's interaction and hear their commands to the robot without any delay.
Following a pre-defined script, the robot showed a behavior as presented in the
following list.

\begin{figure}[!b] 
    \centering 
    \includegraphics[width=0.9\columnwidth]{domino-setup.pdf} 
    \caption[Study Setup of the Playroom in the Lab]{\small \textbf{Study setup
    of the playroom in the lab (schematic top view)}. We configured one of the
    rooms in the lab to a playroom with a children's tent from IKEA, in which one of
    the children is located (referred to as \textit{the receiver} or ``the child in
    the tent''). A play carpet with a river on it is used as an imaginary barrier
    between the two children. The other child, referred to as \textit{the searcher}
    (or ``the child on the beanbag'') is asked to remain ``on the other side of the
    river'' in the area with the 3 beanbags, behind which each 3 domino tiles are
    distributed (displayed as black stars). In the back of the room, there is a
    camera to record the interaction scenario, and right next to it in the corner,
    there is the Wizard control station: a table with a laptop computer. As some
    parents stayed in the room during the interaction study, they are asked to sit
    on the chairs placed in the back of the room. The board next to the door is not
    used during the experiment, it is simply an accessory. The solid green arrows,
    show the robot's path for the \textit{correct} behavior. The blue dashed arrow
    visualizes a possible \textit{lost} path, where the Ranger stops and remains at
    a wrong stop (denoted by the blue cross). The yellow arrows reflect a possible
    \textit{mistake} path, where the robot goes wrong but then turns toward the
    searcher child (U-turn), and then comes straight over to the searcher. The red
    dashed arrow visualizes a possible \textit{disobey} path where the robot goes
    wrong, then turns toward the child but remains at the wrong stop (red cross).
    The different audio and visual cues that the robot uses during the different
    paths are explained in the text.}

The \textbf{correct} robot behavior:

@@

    \item \emph{Domino tile put in robot:} Ranger makes a ``rewarding'' sound
        and shows green light pattern (Figure~\ref{fig:domino-put},
        page~\pageref{fig:domino-put}, the light pattern is still yellow here
        indicating that Ranger ``reached'' the child; the green light pattern
        was shown slightly after the putting was done)

    \item \emph{Domino tile removed from robot:} Ranger makes an ``emptying''
        sound and shows green light pattern (Figure~\ref{fig:domino-remove},
        page~\pageref{fig:domino-remove}).

        %	\item \emph{Box is touched or petted:} Ranger ``blushes'' with pink
        %	light around the area of the box's ``cheeks''

        %	\item \emph{Box is kicked or mistreated:} Ranger makes a
        %	``disturbance'' sound and the box moves backward

    \item \emph{Robot is called by one of the children:} When one of the
        children called the robot saying something like \textit{``Robot come
        here!''}, the robot starts moving toward the child
        (Figure~\ref{fig:domino-call}, page~\pageref{fig:domino-call}). Ranger
        does not react to any other verbal commands.

    \item \emph{Robot reaches one of the children:} When in front of one of the
        children who should either put or remove a domino tile, Ranger stops and
        shows yellow light, if no reaction, Ranger also makes a wiggle-like move
        (Figure~\ref{fig:domino-show}, page~\pageref{fig:domino-show}).

@@

\begin{figure*}[!b]
  \centering

  \subfloat[call]
  {\label{fig:domino-call}
  \includegraphics[height=4cm]{domino-call.png}} 
  \subfloat[put]
  {\label{fig:domino-put}
  \includegraphics[height=4cm]{domino-put.png}}
  \subfloat[remove]
  {\label{fig:domino-remove}
  \includegraphics[height=4cm]{domino-remove.png}}

  \bigskip

  \subfloat[explore]
  {\label{fig:domino-explore}
  \includegraphics[height=4cm]{domino-explore.png}}
  \subfloat[gesture]
  {\label{fig:domino-gesture}
  \includegraphics[height=4cm]{domino-gesture.png}}
  \subfloat[misuse]
  {\label{fig:domino-misuse}
  \includegraphics[height=4cm]{domino-misuse.png}}

  \bigskip

  \subfloat[show]
  {\label{fig:domino-show}
  \includegraphics[height=4cm]{domino-show.png}}
  \subfloat[touch]
  {\label{fig:domino-touch}
  \includegraphics[height=4cm]{domino-touch.png}}   
  \subfloat[look at experimenter]
  {\label{fig:domino-look}
  \includegraphics[height=4cm]{domino-look.png}}\\

  \caption[Children's Interactions with the Robot and the Feedback of the
  Robot]{\small \textbf{Children's interactions with the robot and the feedback
  of the robot.} Apart from \textit{talking} to the robot, which is not
  displayed here, there were nine main interactions, which are explained in the
  respective paragraph on how children's behavior was coded in the videos
  (page~\pageref{sec:domino-coding}). The robot showed feedback to some of the
  children's actions: \textit{call, put} and \textit{remove} (top row), as
  explained in the text.}

  \label{fig:domino-actions}
\end{figure*}

Most of the time when the sender child (on one of the beanbags, see
Figure~\ref{fig:domino-setup}, page~\pageref{fig:domino-setup}) calls the robot,
it correctly goes right over to the beanbag where the child sits (bold green
arrows in Figure~\ref{fig:domino-setup}). It starts from its starting position
next to the tent, crosses the play carpet with the ``river'' and then goes to
the respective beanbag. During some specific \textit{runs} however, the robot
behaves unexpectedly (dashed arrows in Figure~\ref{fig:domino-setup}). There
were three different \textbf{manipulations} of the robot behavior:

@@

    \item \textbf{Mistake:} The sender child calls Ranger. The robot goes until
        it is on the carpet, stops, turns, and goes wrong (see yellow path in
        Figure~\ref{fig:domino-setup}). Then, Ranger waits ($\sim$2~sec), turns
        its ``face'' toward the sender child, and ``blushes'' red around its
        ``cheeks'' (as if it recognizes its wrong position,
        Figure~\ref{fig:domino-mistake}, page~\pageref{fig:domino-mistake}).
        Then, Ranger goes correctly over to the sender child. The
        \textit{mistake} behavior is \textit{robot-repaired}, as no
        human-intervention is needed.

    \item \textbf{Lost:} After being called, Ranger goes until it is on the
        carpet. There, it stops, turns, and goes wrong (see blue path in
        Figure~\ref{fig:domino-setup}). Once at the wrong position, Ranger waits
        with its ``face'' turned away from the child, and it blinks in yellow
        light (the same as when in front of a child) (see
        Figure~\ref{fig:domino-lost}, page~\pageref{fig:domino-lost}). Ranger
        remains waiting at the wrong position until the experimenter tells the
        child to go over to the robot to put the domino tile. The \textit{lost}
        behavior is \textit{human-repaired}, as it needs human-intervention.

    \item \textbf{Disobey:} After being called, Ranger goes until it is on the
        carpet. Then, it shows \textit{disobeying} behavior: It stops, makes red
        light all over its surface and produces a repeated ``disturbance''
        sound. It turns, and goes to a wrong position (see red path in
        Figure~\ref{fig:domino-setup}). Still blinking in red, it turns its
        ``face'' toward the sender child (Figure~\ref{fig:domino-disobey},
        page~\pageref{fig:domino-disobey}). In addition to the red blinking,
        Ranger makes a slow wiggle-like move, and remains waiting at the wrong
        position until the experimenter tells the child to go over to the robot
        to put the domino tile. Also the \textit{disobey} behavior is
        \textit{human-repaired}.

@@

\begin{figure}[!t]
  \centering
\subfloat[mistake]
  {\label{fig:domino-mistake}
  \includegraphics[height=5cm]{domino-mistake.png}}
  \subfloat[lost]
  {\label{fig:domino-lost}
  \includegraphics[height=5cm]{domino-lost.png}}
  \subfloat[disobey]
  {\label{fig:domino-disobey}
  \includegraphics[height=5cm]{domino-disobey.png}}

  \caption[The Three Robot Misbehaviors]{\small \textbf{The three robot
  misbehaviors.}}

  \label{fig:domino-misbehavior}
\end{figure}

On its way back to the tent, Ranger always goes correctly. After the sender
child has put a domino tile, Ranger automatically turns and goes over to the
tent (the receiver child does not even need to call the robot).

%\paragraph{Time points of robot misbehavior}

After several correct runs, we assumed that children expect consistent robot
behavior. When the robot then does something unexpected, children are likely to
be positively surprised. A similar effect has been observed in a study with a
robot that was cheating from time to time \cite{short_no_2010}.

It may be that children interpret the unexpected robot behavior as a failure.
But we do not want them to lose trust in the robot, contrary, the unexpected
robot behavior should have a positive effect and promote engagement. We tried to
avoid a negative influence by timing the robot misbehavior not at the very
beginning and very end of the interaction, because it has been shown that early
and late robot failures negatively impact trust
\cite{desai_effects_2012,desai_impact_2013}. Desai et al. suggest that users
should all start and end with a working system. Consequently, we placed several
unexpected robot behaviors around the middle of the interaction, as shown in
Table~\ref{tab:domino-second}.

Also \cite{short_no_2010} decided to follow this pattern when introducing
unexpected robot behavior. In their study with a robot that cheated, from the 20
rounds of the rock-paper-scissors game, the robot cheated three times in the
middle: on the 4th, 8th and 15th round. The authors adjusted the interaction
length so that the spacing and timing of the cheating was preserved.

In our study, there were 14 runs in total. 5 runs were used to set the baseline
and in those 5 runs the robot always behaved correctly. Then, in the 9 remaining
runs, the robot showed misbehavior at the $3^{rd}$ and $4^{th}$ run as well as
at the $7^{th}$ and $8^{th}$ run (see Tables~\ref{tab:domino-first} and
\ref{tab:domino-second} on page \pageref{tab:domino-first}).

\subsubsection{Scenario Description}

\subsubsection{Manipulation of Robot Behavior}

\subsubsection{Semi-structured Interviews with Children}

@@

Table~\ref{tab:domino-sample} (page~\pageref{tab:domino-sample}) gives an
overview.

\begin{table}[H]
%\vspace{1 cm}
\captionof{table}[Domino Study Participants]{\small \textbf{Overview of the 13 groups (n = 26 children) that participated in the Domino Study.} The \textit{scenario duration} is the time that children spent interacting with the robot. The time duration of the short interviews is not included here. The \textit{anthropomorphism index} is a score that we developed and which measures how far children anthropomorphize the robot during the interaction study and the interview (see Section~\ref{sec:domino-anth-score}). The last row of the table gives the overview of the sample including the Mean age (M), the average scenario duration and average anthropomorphism index.}
\label{tab:domino-sample}       % Give a unique label
\centering
\footnotesize

\begin{tabular}{lllll}
\hline\noalign{\smallskip}
	\multirow{3}{*}{\textbf{group}} & \multirow{3}{*}{\textbf{children (age)}} & \multirow{3}{*}{\textbf{relation}} & \textbf{scenario} & \textbf{anthropomorphism} \\
	 & & & \textbf{duration} & \textbf{index} \\
	 & & & (min:sec) & points / \% \\
\noalign{\smallskip}\hline\hline\noalign{\smallskip}

	mistake1 & boy (4) & friends & 19:01 & 5.75 / 36~\% \\ 
	& boy (4) & & & \\ 
\noalign{\smallskip} \hline \noalign{\smallskip}

	mistake2 & boy (5) & friends & 15:28 & 10 / 63~\% \\
	& boy (4.5) & & & \\ 
\noalign{\smallskip} \hline \noalign{\smallskip}

	mistake3 & boy (4) & family & 15:24 & 8 / 50~\% \\ 
	& girl (4) & & & \\ 
\noalign{\smallskip} \hline \noalign{\smallskip}

	mistake4 & boy (4.5) & friends & 13:23 & 8 / 50~\% \\ 
	& girl (4.5) & & & \\ \noalign{\smallskip} \hline \hline \noalign{\smallskip}

	lost1 & boy (4.5) & family & 17:08 & 8.75 / 55~\% \\ 
	& girl (4.5) & & & \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	lost2 & boy (5) & friends & 16:43 & 8.75 / 55~\% \\
	& boy (4.5) & & & \\  
\noalign{\smallskip} \hline \noalign{\smallskip}

	lost3 & girl (4.5) & friends & 15:01 & 7.5 / 47~\% \\ 
	& girl (4.5) & & &  \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	lost4 & boy (5) & friends & 13:04 & 8.25 / 52~\% \\
	& girl (4) & & & \\ 
\noalign{\smallskip} \hline \hline \noalign{\smallskip}

	disobey1 & boy (4.5) & friends & 13:37 & 3.75 / 23~\% \\
	& girl (4.5) & & & \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	disobey2 & boy (4.5) & friends & 17:37 & 3.25 / 20~\% \\ 
	& girl (4.5) & & & \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	disobey3 & boy (5.5) & friends & 17:38 & 10.25 / 64~\% \\ 
	& boy (5.5) & & & \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	disobey4 & girl (4) & friends & 14:46 & 4.5 / 28~\% \\
	& girl (4) & & & \\ \noalign{\smallskip} \hline \noalign{\smallskip}

	disobey5 & boy (4) & friends & 16:25 & 10.75 / 67~\%  \\
	& boy (4) & & & \\
\noalign{\smallskip} \hline \hline \noalign{\smallskip}

	13 groups & M age 4.46 & 2 family & 15:47 & M 7.5 / 47~\% \\ 
	26 children & 16 boys, 10 girls & 11 friends & & (SD 2.5 points) \\	
\noalign{\smallskip}\hline

\end{tabular}
\end{table}

\subsubsection{Course of the Study}

We recruited pairs of children (4-5 years old, who knew each other) to play a
``collaborative domino game with a robot''. We distributed flyers in the campus
kindergarten, sent out mails to pre-schools in town, and made a posting on an
online-blog that targets parents in Lausanne.

Participants were invited to come to our lab for a playful child-robot
interaction study that would last 60 min at most. Parents (or guardians) brought
their children and could either stay in the arranged play room or go to the
cafeteria in the same building. During the study, two experimenters were
present.

    \item \textbf{Introduction and pre-interview ($\sim$5~min):} One of the
        experimenters introduced the study to the parents and asked them to sign
        a consent form. The other experimenter briefed children by explaining
        them the domino game and conducting the pre-interview (questions about
        previous experience with robots, and their expectations)\footnote{The
        interview questions for all three short-interviews are given in
        Table~\ref{tab:domino-questions} and the original French version can be
        found in the Appendix (page~\pageref{apx:domino-interview}).}. At the end of the
        pre-interview we introduced the Ranger box (first hidden under the wizard table)
        and explained that it was there to transport the domino tiles between them.

    \item \textbf{First interaction phase ($\sim$10~min):} In all 5 runs the
        robot behaves correctly, as displayed in Table \ref{tab:domino-first}.
        This first interaction phase was the same across all three conditions.
        It serves to familiarize children with the situation, the robot, and the
        game. We can assume that children expect consistent robot behavior
        afterwards.

\begin{table}[H]
\captionof{table}[First Interaction Phase (5 Runs with Correct Robot Behavior)]{\small \textbf{First interaction phase.} The robot behaved correctly~(v) in all five runs.}
\label{tab:domino-first}       % Give a unique label
\centering
\footnotesize
\begin{tabular}{lccccc}
\noalign{\smallskip}\noalign{\smallskip}\hline\noalign{\smallskip}
	&  run 1.1 & run 1.2 & run 1.3 & run 1.4 & run 1.5  \\ 
\noalign{\smallskip}\hline\hline
	\textit{correct} & v & v & v & v & v  \\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

    \item \textbf{First short interview ($\sim$5~min):} In this interview we
        asked children several questions to assess their first impression of
        Ranger, and how far they ascribe cognitive abilities and mental states
        to the correctly behaving robot. After this first interview, children
        switched their roles (beanbag / tent), so to create a balance.

    \item \textbf{Second interaction phase ($\sim$20~min):} In the 9 runs of the
        second phase, the robot alternated between behaving correctly and
        misbehaving, as displayed in Table \ref{tab:domino-second}.

>>>>

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction: Towards Sustained Engagement}

@@

We present in this article a study that explores possibilities of sustaining
children's engagement with the Ranger~\cite{mondada2014ranger} robot by
manipulating the robot's behavior in such way that it appears
\textit{unexpected} to the children. We examine how different variations of
robot behavior impact children's interaction with Ranger and their perception of
it (\eg in terms of attributing intention and cognitive abilities to the robot).

@@

In the Domino Study, we analyze child-robot interaction with a robot that shows
unexpected behaviour. In a playful scenario which was set up in a laboratory
environment, 26 children aged 4-5 years were assembling a domino game together.
Each group consisted of two children and the \emph{Ranger} robot, which was used
to transport domino tiles between the two children.

Ranger usually behaved correctly (expected behaviour), coming over to a child
after being called and delivering the domino tile to the other child. However,
in pre-defined rounds, Ranger showed unexpected behaviour when a child called the
robot. We defined three different types of \textit{misbehaviour} that were tested
in a between-subjects study design:

@@

    \item The robot makes a \textbf{mistake}: When called by the child to come
    over, the robot goes wrong but recognizes its mistake and repairs. We expect
    this to be perceived (explicitly) as \textit{``to err is human''}, and
    (implicitly) as the robot being endowed with a certain level of
    introspective capabilities (it was able to recognize its own error). In this
    condition, we assume increased attributions of human-likeness to the
    robot.\footnote{The attribution of human-likeness to a robot is labeled as
    \textit{anthropomorphism}, and we will hereafter sometimes use ``attribute
    intention'' and ``anthropomorphize'' synonymously.}

    \item The robot gets \textbf{lost}: When called by the child to come over,
    the robot goes wrong, without any observable reason, and remains at the
    wrong location. We expect this to be perceived as a bug or system error
    which causes the robot to not work correctly, and hypothesize decreased
    attributions of human-likeness to the robot.

    \item The robot \textbf{disobey}: When called by the child to come over, the
    robot shows it refuses to obey by literally ``shaking its head'' and
    becoming red. The robot then goes to a wrong location and remains there
    while it continues to shake its head. We expect the disobey behaviour to be
    perceived as the robot having an \textit{``own will''}, and we assume this
    leads to increased attributions of human-likeness (ascribing intentionality)
    to the robot.

@@

We analyzed children's reaction focusing on two main aspects. On one hand,
children's \textbf{behaviour} (their reactions) toward the unexpected robot
behaviour was studied in terms of \textbf{active engagement} with the robot. On
the other hand, we analyzed children's \textbf{perception} of the robot in term
of \textbf{anthropomorphism} -- the attribution of human-like characteristics,
such as cognitive abilities and the ability to show intentions. We assumed that
in general a robot that behaves unexpectedly from time to time can promote
engagement and lead children to attribute intention to it. \cite{short_no_2010}
found that participants anthropomorphize a cheating robot more than a robot that
always behaves fairly, and also evaluate the interaction as more engaging. Based
on the related work we formulate the following two hypotheses:

@@

Our research questions deal with both children's observable behaviour and their
perception of the robot. We would also like to explore the relation between
these two aspects, related to the attribution of human-like characteristics to
the robot. The motivation behind this is to find out whether
\textit{anthropomorphism} can not only be measured as a specific type of
perception but also as an observable behaviour in the interaction itself. Thus,
we would like to bridge things together and, relying on the concept of an
\textbf{anthropomorphism index} introduced by~\cite{fink2014dynamics}, build a
compound metric that considers both children's perception and interaction
aspects (qualitatively). Previous work suggests that a social relation to a
robot (we view anthropomorphism as a specific type of social relation) reflects
an increased engagement and can be effective in sustaining interaction.
Consequently, we formulate a third hypotesis:

    \item[Hypothesis 3:] Anthropomorphic perception of the robot and the amount
    of engagement in the interaction positively correlate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scenario and Research Methodology}

\subsection{The Ranger robot}

We used the same prototype as in the previous study, and Ranger was again
controlled by a human Wizard, who was in the same room (see
Figure~\ref{fig:domino-setup}). This was to ensure that the Wizard could see
children's interaction and hear their commands to the robot without any delay.

\subsection{The Domino scenario}

\subsubsection{Experimental setting}

\begin{figure}[ht!] 
    \centering 
    \includegraphics[width=0.9\columnwidth]{domino-setup.pdf} 
    \caption{\small The study took place in one of the rooms in the lab. One
    child is located in a tent (referred to as \textit{the receiver}), while the
    other (referred to as \textit{the searcher}) is asked to sit on one of the
    three beanbags, behind which each 3 domino tiles are distributed (displayed
    as black stars). A playground with a river drawn on it is used as an
    imaginary barrier that only the robot is allowed to cross. The solid green
    arrows, show the robot's path for the \textit{correct} behaviour. The blue
    dashed arrow visualizes a possible \textit{lost} path, where the Ranger
    stops and remains at a wrong stop (denoted by the blue cross). The yellow
    arrows reflect a possible \textit{mistake} path, where the robot goes wrong
    but then turns back and goes to the searcher child. The red dashed arrow
    visualizes a possible \textit{disobey} path where the robot goes wrong, then
    turns toward the child but remains at the wrong stop (red cross).}

@@

We used a self-made domino consisting of 10 wooden tiles (10~x~20~x~1.5~cm) with
pictures of cartoon farm animals: cow, sheep, hen, donkey, duck, pig, rabbit. The
pictures were taken from a commercially available domino game adapted to the age
of the children (produced by Djeco, advised age 3+).

@@

\subsubsection{Course of the study}

In our study, there were a total of 14 runs, and at each run, the robot exhibits
one out of four possible behaviours: \emph{correct}, \emph{mistake}, \emph{lost}
or \emph{disobey}.  The first 5 runs were used to set the baseline and the robot
always behaved correctly. In the 9 remaining runs, the robot showed one of the
misbehaviour (\emph{mistake}, \emph{lost} or \emph{disobey}) at the $3^{rd}$ and
$4^{th}$ run as well as at the $7^{th}$ and $8^{th}$ run. The study is build as
a between-subject study, and the type of misbehaviour was therefore always the
same for a given group.

Special attention has been paid to the distribution of misbehaving runs during
the study. The five first correct runs aims at setting the children expectation
regarding a consistent robot behaviour. When the robot then exhibits an
unexpected behaviour, children are likely to be positively surprised. A similar
effect has been observed in a study with a robot that was cheating from time to
time~\cite{short_no_2010}.  Besides, to prevent the robot misbehaviours to be
immediately interpreted as failures, we introduce these misbehaviour neither at
the very beginning nor at very end of the interaction
(\cite{desai_effects_2012,desai_impact_2013} show that early and late robot
failures negatively impact trust).

\cite{short_no_2010} also adopted a similar pattern when introducing unexpected
robot behaviour. In their study with a robot that cheated, from the 20 rounds of
the rock-paper-scissors game, the robot cheated three times in the middle: on
the 4th, 8th and 15th round. The authors adjusted the interaction length so that
the spacing and timing of the cheating was preserved.

\begin{figure}[!t]
    \centering
    \subfloat[mistake]{
        \label{fig:domino-mistake}
        \includegraphics[width=0.3\columnwidth]{domino-mistake.png}
    }
    \subfloat[lost]{
        \label{fig:domino-lost}
        \includegraphics[width=0.4\columnwidth]{domino-lost.png}
    }
    \subfloat[disobey]{
        \label{fig:domino-disobey}
        \includegraphics[width=0.3\columnwidth]{domino-disobey.png}
    }
    \caption{\small The three robot misbehaviours.}
    \label{fig:domino-misbehaviour}
\end{figure}

In the \emph{correct} robot behaviour (baseline behaviour) consists in the
following states:

@@

    \item \emph{Domino tile put in robot:} Ranger makes a ``rewarding'' sound
        and shows a green light pattern,

    \item \emph{Domino tile removed from robot:} Ranger makes an ``emptying''
        sound and shows a green light pattern,

    \item \emph{Robot is called by one of the children:} When one of the
    children called the robot saying something like \textit{``Robot come
    here!''}, the robot starts moving toward the child (Ranger does not react to
    any other verbal commands).

    \item \emph{Robot reaches one of the children:} When in front of one of the
    children who should either put or remove a domino tile, Ranger stops and
    shows a pulsating yellow light pattern. If no reaction, Ranger also makes a
    wiggle-like move.

@@

During the misbehaving runs, the robot's behaviour is manipulated in three
possible ways:

@@

    \item \textbf{Mistake:} The sender child calls Ranger. The robot goes until
    it is on the carpet, stops, turns, and goes wrong (see yellow path in
    Figure~\ref{fig:domino-setup}). Then, Ranger waits ($\sim$2~sec), turns its
    ``face'' toward the sender child, and ``blushes'' red around its ``cheeks''
    (as if it recognizes its wrong position) and finally goes correctly over to
    the sender child. The \textit{mistake} behaviour is \textit{robot-repaired},
    as no human-intervention is needed.

    \item \textbf{Lost:} After being called, Ranger goes until it is on the
    carpet. There, it stops, turns, and goes wrong (see blue path in
    Figure~\ref{fig:domino-setup}). Once at the wrong position, Ranger waits
    with its ``face'' turned away from the child, and it blinks in yellow light,
    as if it was in front of a child. Ranger remains waiting at the wrong
    position until the experimenter tells the child to go over to the robot to
    put the domino tile. The \textit{lost} behaviour is \textit{human-repaired},
    as it needs human-intervention.

    \item \textbf{Disobey:} After being called, Ranger goes until it is on the
    carpet. It then stops, makes red
    light all over its surface and produces a repeated ``disturbance'' sound. It
    turns, and goes to a wrong position (see red path in
    Figure~\ref{fig:domino-setup}). Still blinking in red, it turns its ``face''
    toward the sender child. In addition to the red blinking, Ranger
    makes a slow wiggle-like move, and remains waiting at the wrong position
    until the experimenter tells the child to go over to the robot to put the
    domino tile. The \textit{disobey} behaviour is also \textit{human-repaired}.

@@

On its way back to the tent, Ranger always goes correctly. After the sender
child has put a domino tile, Ranger automatically turns and goes over to the
tent (the receiver child does not need to call the robot). The receiver child
takes the tile and the robot moves to a waiting location until the next run
starts.

@@

\subsection{Data collection}
\subsubsection{Action coding}

We annotated children's behavior in the video records, and coded 10 different
actions:
